VFTS682: a confirmed dynamical ejection?

### 2018-05-10 VFTS682.tex
new fig. wrote something for kinematic age

Using the 651 stars selected as described in \Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle =
0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
in good agreement with what found by \cite{lennon:18}. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$.
These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=38\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=26\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

@@

% The inset plot in \Figref{fig:main} shows a zoom in of the region
% around R136 which includes VFTS682. Stars belonging to our ``R136
% cluster'' group are shown in green (see \Secref{sec:sample}), together with their proper motion
% as listed in the Gaia DR2 catalog.
% To test the hypothesis that VFTS682 was indeed ejected from this region, we check the
% orientation of its proper motion relative to these stars. We further
% consider also the kinematic age of the star.

% We begin by evaluating the average proper motion for the stars
% representing the cluster. This is necessary to constrain the actual velocity
% obtained by VFTS682 at its ejection. By using \Eqref{eq:mean} we find  
% $\langle\mu_\mathrm{RA}^{R136}\rangle = 1.750\pm0.008\,\mathrm{mas\ yr^{-1}}$
% and $\langle\mu_\mathrm{DE}^\mathrm{R136}\rangle =
% 0.742\pm0.009\,\mathrm{mas\ yr^{-1}}$. Therefore, using the values
% from \Tabref{tab:vfts682}, we obtain the components of the proper
% motion for VFTS682 relative to R136 $\mu^{R136}_\mathrm{RA} = 0.09\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^{R136}_\mathrm{DE} =
% 0.04\pm 0.08\,\mathrm{mas\ yr^{-1}}$.\todo{improve numbers by
%   selecting better sample}.
% \todo{discuss if direction is compatible
% The error cone, obtained by
% propagating the errors on the proper motion components indicates that
% R136 is indeed the most likely origin, as expected by \cite{fujii:11,
%   banerjee:12}.
% }

% Assuming again a distance of 50\,kpc and
% not considering the error on the distance, we can convert these into
% the components of the tangential velocity relative to R136,
% $v^\mathrm{R136}_\mathrm{RA}=21 \pm 18\, \kms$ and $v^\mathrm{R136}_\mathrm{DE}=10
% \pm 21\,\kms$. Finally, using the peculiar line of sight velocity
% from \cite{bestenlehner:11}, we obtain the three-dimensional peculiar
% velocity of VFTS682 with respect to R136

% \begin{equation}
%   \label{eq:speed_R136}
%   v^\mathrm{sur} = \sqrt{\left(v^{R136}_\mathrm{RA}\right)^2
%     +\left(v^{R136}_\mathrm{DE}\right)^2+\left(\delta v_\mathrm{rad}\right)^2} = 51 \pm 23
%   \, \kms \ .
% \end{equation}

\subsection{Kinematic age}

@@

\begin{acknowledgements}
  The background image of \Figref{fig:main} is based on observations
  made with ESO Telescopes at the La Silla Observatory under programme
  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
  This research made use of APLpy, an open-source plotting package for Python \cite{robitaille:12}.
  We are grateful to S.~Torres for extremely useful discussions on the
  Gaia DR2 dataset.
\end{acknowledgements}
\end{document}

>>>>

Using the 651 stars selected as described in \Secref{sec:sample} and shown in blue in \Figref{fig:main}, we find averaged proper motion components of
$\langle\mu_\mathrm{RA}^\mathrm{sur}\rangle = 1.683\pm0.002\,\mathrm{mas\ yr^{-1}}$ and
$\langle\mu_\mathrm{DE}^\mathrm{sur}\rangle =
0.672\pm0.003\,\mathrm{mas\ yr^{-1}}$. We note that these values are
in good agreement with what found by \cite{lennon:18}. Subtracting these values from the
proper motions of VFTS682 (see \Tabref{tab:vfts682}), we obtain the
components of proper motion of the star relative to the surrounding region
$\mu^\mathrm{sur}_\mathrm{RA} = 0.16\pm 0.07\,\mathrm{mas\ yr^{-1}}$ and $\mu^\mathrm{sur}_\mathrm{DE} =
0.11\pm 0.08\,\mathrm{mas\ yr^{-1}}$. We note that the error budget is
dominated by the errors on the proper motion components of VFTS682.

These can be converted in the
the components of the transverse velocity $v^\mathrm{sur}_\mathrm{RA}=38\pm17\,\kms$,
$v^\mathrm{sur}_\mathrm{DE}=26\pm19\,\kms$, assuming a distance of
50\,kpc (we do not account for the uncertainty in the distance
estimate when propagating errors). The radial velocity from
\cite{bestenlehner:11} then gives the third component along
the line of sight, allowing us to calculate the speed of the star:

@@

The red arrow in \Figref{fig:main} shows the proper motion of VFTS682
relative to the region, and the lighter red arrows show the possible
range of directions within the uncertainties in the measured proper
motion. It is clear that the most likely origin of the star is R136,
as predicted by \cite{fujii:11, banerjee:12}.

%\subsection{Kinematic age}

We can therefore consider the kinematic age for this star assuming
that it originates from the cluster,

\begin{equation}
  \label{eq:kin_age}
  \tau_\mathrm{kin} = \frac{d_\parallel}{v_\parallel} =
  \frac{29\,\mathrm{pc}}{46\,\kms} \simeq 0.63\, \mathrm{Myr} \ \ ,
\end{equation}
where $d_\parallel =29$\,pc is the projected distance from VFTS682 to
the core of R136 \citep[][]{bestenlehner:11}, $v_\parallel = 46\pm
17\,\kms$ is the speed of the star relative to the cluster projected on the sky, obtained
using the relative proper motion components calculated in
\Secref{sec:runaway}, and we use the approximation $1 \kms \simeq 1
\mathrm{pc \ Myr^{-1}}$.

@@

\begin{acknowledgements}
  The background image of \Figref{fig:main} is based on observations
  made with ESO Telescopes at the La Silla Observatory under programme
  ID 076.C-0888, processed and released by the ESO VOS/ADP group.
  This research made use of APLpy, an open-source plotting package for Python \cite{robitaille:12}.
  We are grateful to S.~Torres for extremely useful discussions on the
  Gaia DR2 dataset, and to M.~C.~Ramirez-Tannus for help with APLpy.
\end{acknowledgements}
\end{document}
